---
title: "Problem set 1 Siyuan Wu"
output: html_document
date: "2024-08-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

## Problem Set 1

### Problem 1 Wine Data

#### a

```{r}
wine <- read.csv("~/Downloads/umich/2024 fall/stats 506/hw/problem set1/wine/wine.data", header=FALSE)
colnames(wine)<-c("class","Alcohol", "Malic_acid",
 	"Ash",
	"Alcalinity_of_ash",  
 	"Magnesium",
	"Total_phenols",
 	"Flavanoids",
 	"Nonflavanoid_phenols",
 	"Proanthocyanins",
	"Color_intensity",
 	"Hue",
 	"OD280/OD315_of_diluted_wines",
 	"Proline")
```

#### b

```{r}
table(wine$class)
```

It is same as the document noted about the numbers of wine in each class.

#### c

##### 1

```{r}
cor1<-cor(wine$Alcohol,wine$Color_intensity)
```

The correlation is `r cor1`.

##### 2

```{r}
wineclass1<-wine[which(wine$class==1),]
wineclass2<-wine[which(wine$class==2),]
wineclass3<-wine[which(wine$class==3),]
cor2_1<-cor(wineclass1$Alcohol,wineclass1$Color_intensity)
cor2_2<-cor(wineclass2$Alcohol,wineclass2$Color_intensity)
cor2_3<-cor(wineclass3$Alcohol,wineclass3$Color_intensity)
print(c(cor2_1,cor2_2,cor2_3))
```

The correlation of class 1 is highest, while correlation of class 2 is lowest.

##### 3

```{r}
targetac<-wine[which.max(wine$Color_intensity),]$Alcohol
```

The alcohol content is `r targetac`.

##### 4

```{r}
temp=0
for (i in 1:nrow(wine)){
  if (wine[i,]$Proanthocyanins>wine[i,]$Ash){
    temp=temp+1
  }
}
perc<-temp/nrow(wine)*100
```

The percentage is `r perc`%.

#### d

```{r}

tableD<-t(data.frame(colMeans(wine),colMeans(wineclass1),colMeans(wineclass2),colMeans(wineclass3)))
rownames(tableD)<-c("All","class_1","class_2","class_3")
tableD<-tableD[,-1]
tableD
```

#### e

```{r}

t1=wine[which(wine$class==1|wine$class==2),]
t.test(Total_phenols~ class, t1)
t2=wine[which(wine$class==2|wine$class==3),]
t.test(Total_phenols~ class, t2)
t3=wine[which(wine$class==1|wine$class==3),]
t.test(Total_phenols~ class, t3)
```

Based on the result that all the p-values are smaller than 0.05, we can have the conclusion as the expectation of level of phenols differs across the three classes.

### Problem 2 **AskAManager.org Data**

#### a

```{r}
AskAManager<-read.csv("~/Downloads/umich/2024 fall/stats 506/hw/problem set1/AskAManager.csv")
```

#### b

```{r}
colnames(AskAManager)<-c("Survey_Id","Time","Age_Range","Industry","Job","Job_Additional_Context","Annual_Salary","Additional_Bonus","Salary_Currency","Other_Salary_Currency","Salary_Additional_Context","Country_of_Work","US_state_of_Work","City_of_Work","Years_of_Work","Years_of_Work_in_Field","Education_Level","Gender","Race")
```

#### c

```{r}
orsampsize=nrow(AskAManager)
AskAManagerUSD=AskAManager[which(AskAManager$Salary_Currency=="USD"|AskAManager$Other_Salary_Currency=="USD"),]
sampsizec=nrow(AskAManagerUSD)
```

The sample size is decreased from `r orsampsize` to `r sampsizec`.

#### d

```{r}
#unique(AskAManagerUSD$Age_Range)
#unique(AskAManagerUSD$Years_of_Work)
#unique(AskAManagerUSD$Years_of_Work_in_Field)
AskAManagerUSD$Work_low_bond=case_when(AskAManagerUSD$Years_of_Work=="5-7 years"~5, 
AskAManagerUSD$Years_of_Work=="2 - 4 years"~2, 
AskAManagerUSD$Years_of_Work=="8 - 10 years"~8,  
AskAManagerUSD$Years_of_Work=="21 - 30 years"~21,         
AskAManagerUSD$Years_of_Work=="11 - 20 years"~11, 
AskAManagerUSD$Years_of_Work=="41 years or more"~41,
AskAManagerUSD$Years_of_Work=="31 - 40 years"~31,
AskAManagerUSD$Years_of_Work=="1 year or less"~0)

AskAManagerUSD$Work_low_bond_Field=case_when(AskAManagerUSD$Years_of_Work_in_Field=="5-7 years"~5, AskAManagerUSD$Years_of_Work_in_Field=="2 - 4 years"~2, AskAManagerUSD$Years_of_Work_in_Field=="8 - 10 years"~8,  AskAManagerUSD$Years_of_Work_in_Field=="21 - 30 years"~21,          AskAManagerUSD$Years_of_Work_in_Field=="11 - 20 years"~11, AskAManagerUSD$Years_of_Work_in_Field=="41 years or more"~41,AskAManagerUSD$Years_of_Work_in_Field=="31 - 40 years"~31,AskAManagerUSD$Years_of_Work_in_Field=="1 year or less"~0)

AskAManagerUSD$Age_high_bond=case_when(
  AskAManagerUSD$Age_Range=="25-34"~34,
   AskAManagerUSD$Age_Range=="45-54"~54,
   AskAManagerUSD$Age_Range=="35-44"~44,
   AskAManagerUSD$Age_Range=="18-24"~24,
   AskAManagerUSD$Age_Range=="65 or over"~100,#set a maximum here
   AskAManagerUSD$Age_Range=="55-64"~64,
  AskAManagerUSD$Age_Range=="under 18"~18
)

```

```{r}
print(sum(is.na(AskAManagerUSD$Age_high_bond)))
print(sum(is.na(AskAManagerUSD$Work_low_bond)))
print(sum(is.na(AskAManagerUSD$Work_low_bond_Field)))
```

```{r}
AskAManagerUSD=AskAManagerUSD[which(AskAManagerUSD$Work_low_bond>=AskAManagerUSD$Work_low_bond_Field &AskAManagerUSD$Work_low_bond+18<=AskAManagerUSD$Age_high_bond),]
sampsized=nrow(AskAManagerUSD)
```

The sample size is `r sampsized`.

#### e

```{r}
print(sum(is.na(AskAManagerUSD$Annual_Salary)))
boxplot(AskAManagerUSD$Annual_Salary,main="boxplot of annual salary",ylab="annual salary")
hist(AskAManagerUSD[which(AskAManagerUSD$Annual_Salary<500),]$Annual_Salary, main=("hist of annual salary under 500"),xlab="annual salary")
```

From the box plot, we will delete the salary over $10^7$ , as they are far from the other observations. Besides, with the histgram, we can see a lot of low values are between 0 and 200, which will also be deleted.

```{r}
AskAManagerUSD=AskAManagerUSD[which(AskAManagerUSD$Annual_Salary<10^7&AskAManagerUSD$Annual_Salary>200),]
sampsizefinal=nrow(AskAManagerUSD)

```

The final sample size is `r sampsizefinal`.

### Problem 3 **Palindromic Numbers**

#### a

```{r}
isPalindromic<-function(a){
  if(!is.numeric(a)|a<0|!a%%1==0) stop("must be a positive integer")
  temp=a
  t=0
while (temp>0){
  t=t*10+temp%%10
  temp=temp%/%10
}
return(list(isPalindromic=(t==a),reversed=t))
}
```

```{r}
isPalindromic(728827)
isPalindromic(39951)
```

#### b

```{r}
nextPalindrome<-function(a){
  if(!is.numeric(a)|a<0|!a%%1==0) stop("must be a positive integer")
  t=a+1
  while (!isPalindromic(t)$isPalindromic){t=t+1
  }
  return(t)}
```

```{r}
nextPalindrome(7152)
nextPalindrome(765431537)
```

#### c

```{r}
nextPalindrome(391)
nextPalindrome(9928)
nextPalindrome(19272719)
nextPalindrome(109)
nextPalindrome(2)
```
